---
title: "Tweets Sentimental Analysis        -- Thomaskutty_20122011"
output:
  pdf_document: default
  html_notebook: default
---

# Importing the neessary libraries 
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytext)
library(glue)
library(stringr)

library(SnowballC)
library(tm)
library(twitteR)
library(syuzhet)
```


# Reading the file and assigning  proper column names fo the dataframe. 
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
df = read.csv('/cloud/project/tweets.csv', nrows = 50000, header = TRUE, sep = ',')
attach(df)
df = subset(df, select = c(-X0))
colnames(df) = c('id', 'date', 'flag', 'user','tweet')
attach(df)
head(df)

```

# Printing the first 10 tweets
```{r}
for(i in seq(1:10)){
  print(df$tweet[i])
}
```
# Printing the final dataframe of tweets 
```{r paged.print=FALSE}
head(df)
```

# Creating a function to clean the tweets
```{r}
library(tidyverse)  
clean_tweets <- function(x) {
            x %>%
                    str_remove_all(" ?(f|ht)(tp)(s?)(://)(.*)[.|/](.*)") %>%
                    str_replace_all("&amp;", "and") %>%
                    str_remove_all("[[:punct:]]") %>%
                    str_remove_all("^RT:? ") %>%
                    str_remove_all("@[[:alnum:]]+") %>%
                    str_remove_all("#[[:alnum:]]+") %>%
                    str_replace_all("\\\n", " ") %>%
                    str_to_lower() %>%
                    str_trim("both")
  return (x)
}
# applying the cleaning function to tweet column 
df$tweet = sapply(df$tweet, clean_tweets)
```

# Getting the sentiments of each tweets
```{r}
get_sents = function(tweet){
  sent.value = get_sentiment(tweet)
  return (sent.value) 
}

# applying the get_sents function to tweet column 
df$sentiment_score = sapply(df$tweet, get_sents) 
```


# printing the final dataframe 
```{r warning=FALSE, paged.print=FALSE}
head(df)
```
# Getting both negative and positive tweets in a seperate dataframe 
```{r}
positive.tweets = df[df$sentiment_score > 0,]
negative.tweets = df[df$sentiment_score < 0,]
```

# Assigning labels to each tweets 
```{r paged.print=FALSE}
to_sent_labels = function(x){
  if (x >0){
    'positive'
  }else if (x <0){
    'negative'
  }else{
    'neutral'
  }
}

df$sent_labels = sapply(df$sentiment_score , to_sent_labels)
head(df)
```

# Getting the count of sentiment labels 

```{r}
print(table(df$sent_labels))
library(ggplot2)
ggplot(data = df, aes(sent_labels)) + geom_bar(fill = 'lightgreen') + theme_minimal()
```
# Interpretation 

We can see that the counts of both positive and negative tweets are almost same. So tweets are equally distributed ( both positive and negative).











